---
title: "AIDS Drug"
author: "Gracie Telles"
date: "Fall 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
library(dplyr)
druggies = filter(triUmp, triUmp$DrugUser == 'yes')
```

# Introduction

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  

$H_0$ There is no deifference between the new preventative AIDS drug treatment and the effct it has on users.

$H_A$ There is a difference between the new preventative AIDS drug treatment and the effect it has on users.


# Methods

Both the response variable and the exploratory variable are categorical, therefore we will us the cat-cat helper file.

# Results

## Descriptive Results


### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=druggies, type="percent")
```

This graph shows that for those people who got infected, a very high amount got the placebo, as compared to those who got the preventative drug. 


### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=druggies)
rowPerc(table1)
```

This table shows the placebo group is about 3 times as likely to get infected as opposed to those that received the preventative drug.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)

```

We reject the null hypothesis because the P value is 0.0000001242 which is below 0.05.

```{r}
druggies$infected = factor(druggies$infected, c('none', 'aids'))
fisher.test(druggies$treatment,druggies$infected)
```
The P value in this test is equal to .0000000423. The people who got the vaccine are almost 3 times more protected from AIDS compared to the placebo receiving people.


# Discussion and Conclusion

The vaccine is effective in preventing AIDS. the graphical data showed the significant difference in the people with AIDS who received the vaccine compared to the placebo vaccine. Our numerical and inferential data also tells us that the vaccine was 3 times more protective to aids than the placebo vaccine. The Chisq test which gave us a P-Value less than 0.05, which means we reject the null hypothesis. In conclusion, the new preventative drug treatment is effective against the infection AIDS. 