---
title: "AIDS Drug"
author: "Gracie Telles"
date: "Fall 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
library(dplyr)
druggies = filter(triUmp, triUmp$DrugUser == 'yes')
```

# Introduction

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  

$H_0$ There is no difference between the new preventative AIDS drug treatment and the effect it has on users.

$H_A$ There is a difference between the new preventative AIDS drug treatment and the effect it has on users.


# Methods

Both the response variable and the exploratory variable are categorical, therefore we will us the cat-cat helper file.

# Results

## Descriptive Results


### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=druggies, type="percent")
```

This graph shows that for those people who got infected, the majority received the placebo, as compared to those who received the preventative drug. 


### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=druggies)
rowPerc(table1)
```

This table shows the placebo group is about 3 times as likely to get infected as opposed to those that received the preventative drug. This shows that the active preventative is working.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)

```

We reject the null hypothesis because the P value is 0.0000001242 which is below 0.05.

```{r}
druggies$infected = factor(druggies$infected, c('none', 'aids'))
fisher.test(druggies$treatment,druggies$infected)
```
The P value in this test is equal to .0000000423. The data shows that the people who received the vaccine were three times more protected from AIDs than those who recieved the placebo.


# Discussion and Conclusion

The vaccine is effective in preventing AIDS. The graphical data showed the significant difference in the people with AIDS who received the vaccine compared to the placebo vaccine. Our numerical and inferential data also tells us that the vaccine was three times more protective from AIDs than the placebo. The Chisq test gave us a P-Value less than 0.05, which means we reject the null hypothesis- there IS a very marked difference in those who were given the vaccine as opposed to the placebo. In conclusion, the new preventative drug treatment is effective against the infection AIDS. 